---
title: "R-project-patent"
format: html
---

## quarto document

## Link to the Github Repository

https://github.com/brl-noe/R-project-patent.git

## Loading the library
```{r}
#| message: false
library(here)
library(vroom)
library(dplyr)
library(ggplot2)
library(tidyr)
theme_set(theme_bw())
```


## Loading the data

```{r}
here::i_am("R-project-patent.Rproj")
```

```{r}
## Loading patent data
patent <- vroom(here("patent.tsv"))

## Loading key indicator data
key_indicator <- vroom(here("key_indicator.tsv"))

## Loading education data
education <- vroom(here("education.tsv"))
```

## Basics informations about the dabases
```{r}
# Function to generate a data summary
get_data_summary <- function(df) {
  rows <- nrow(df)
  cols <- ncol(df)
  sample <- head(df)
  
  return(list(
    NumRows = rows,
    NumCols = cols,
    SampleData = sample
  ))
}

# Get summaries for each dataset
patent_summary <- get_data_summary(patent)
key_indicator_summary <- get_data_summary(key_indicator)
education_summary <- get_data_summary(education)

# Combine the summaries into a single data frame
data_summaries <- data.frame(
  Dataset = c("Patent", "Key Indicator", "Education"),
  NumRows = c(patent_summary$NumRows, key_indicator_summary$NumRows, education_summary$NumRows),
  NumCols = c(patent_summary$NumCols, key_indicator_summary$NumCols, education_summary$NumCols),
  SampleData = list(patent_summary$SampleData, key_indicator_summary$SampleData, education_summary$SampleData),
  stringsAsFactors = FALSE
)

# Print the summary data frame
print(data_summaries)

# Display the sample data for each dataset
for (i in 1:nrow(data_summaries)) {
  cat(paste0("Sample data for ", data_summaries$Dataset[i], ":\n"))
  print(data_summaries$SampleData[[i]])
  cat("\n")
}
```

## Description of the data joining, cleaning, etc.
```{r}
# Join the datasets
combined_data <- left_join(education, key_indicator, by = "country")
combined_data <- left_join(combined_data, patent, by = "country")
```

The `left_join()` function from the `dplyr` package will perform a left join between the datasets, using the "country" column as the common key. This will ensure that all rows from the education dataset are included, with the corresponding data from the other two datasets.

Data cleaning and transformation:
```{r}
# Clean and tidy the data
combined_data <- combined_data %>%
  mutate(country = tolower(country)) %>% 
  pivot_longer(cols = starts_with("X"), names_to = "year", values_to = "value") %>%
  mutate(year = as.numeric(str_extract(year, "\\d+")))
```

We convert the country names to lowercase for consistency. Then we reshape the data into a tidy format, with each row representing a country-year observation. Finally we extract the year from the column names and converting it to a numeric format.

